\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{mathtools}
\usepackage{parskip}
\usepackage{hyperref}

\title{Toward the Cs\'asz\'ar Polyhedron via Cell-Complex Enumeration\\[2pt]
\large Step 1: Notation and the First Conditions}
\author{}
\date{}

\begin{document}
\maketitle

\section*{Notation}

Let $P_i = (x_i, y_i, z_i) \in \mathbb{R}^3$ be the coordinate vector of the symbolic
point $i$. Define the \emph{bracket}
\[
[abcd] \;:=\; \det\!\begin{pmatrix} 1 & x_a & y_a & z_a \\ 1 & x_b & y_b & z_b \\ 1 & x_c & y_c & z_c \\ 1 & x_d & y_d & z_d \end{pmatrix}
\;=\; \det\!\bigl(P_b-P_a,\; P_c-P_a,\; P_d-P_a\bigr),
\]
which is $6$ times the signed volume of the tetrahedron $abcd$ and is alternating in
$a,b,c,d$. With this,
\[
abc^{\pm} \;:=\; \bigl\{\,p\in\mathbb{R}^3 \;:\; \pm[a,b,c,p] > 0\,\bigr\}.
\]

\section*{Condition 1: \texorpdfstring{$3 \in 012^+$}{3 in 012+}}

This is just $[0,1,2,3] > 0$. Three equivalent forms:

\paragraph{(a) Determinant.}
\[
\det\!\begin{pmatrix} 1 & x_0 & y_0 & z_0 \\ 1 & x_1 & y_1 & z_1 \\ 1 & x_2 & y_2 & z_2 \\ 1 & x_3 & y_3 & z_3 \end{pmatrix} > 0
\quad\Longleftrightarrow\quad
\det\!\begin{pmatrix} x_1-x_0 & y_1-y_0 & z_1-z_0 \\ x_2-x_0 & y_2-y_0 & z_2-z_0 \\ x_3-x_0 & y_3-y_0 & z_3-z_0 \end{pmatrix} > 0
\]

\paragraph{(b) Scalar triple product (cross-product antisymmetry).}
\[
(P_1 - P_0)\,\cdot\,\bigl((P_2 - P_0)\times(P_3 - P_0)\bigr) \;>\; 0.
\]
This form privileges $P_0$ as a basepoint, hiding the $S_4$-antisymmetry on the
four points. \emph{Fully unparenthesized.} Distribute the differences (using
$a\!\times\!a=0$ and the cyclic identity $a\!\cdot\!(b\!\times\!c)=c\!\cdot\!(a\!\times\!b)$);
the bracket reorganizes into the alternating sum over $3$-element subsets of
$\{0,1,2,3\}$:
\[
[0,1,2,3] \;=\; P_1\!\cdot\!(P_2\!\times\!P_3) \;-\; P_0\!\cdot\!(P_2\!\times\!P_3) \;+\; P_0\!\cdot\!(P_1\!\times\!P_3) \;-\; P_0\!\cdot\!(P_1\!\times\!P_2).
\]
Equivalently, with $T(i,j,k) := P_i\cdot(P_j\!\times\!P_k)$ (cyclically symmetric in
$i,j,k$) and $\widehat{d}$ denoting the elements of $\{0,1,2,3\}\setminus\{d\}$ in
increasing order,
\[
[0,1,2,3] \;=\; \sum_{d=0}^{3} (-1)^d\, T\!\bigl(\widehat{d}\bigr).
\]
The $S_4$-antisymmetry is now manifest --- any swap of two points flips the
bracket's sign.

\paragraph{(c) Polynomial in coordinates.}
With $\Delta_i := P_i - P_0$ for $i=1,2,3$:
\[
\Delta_{1,x}(\Delta_{2,y}\Delta_{3,z} - \Delta_{2,z}\Delta_{3,y})
+ \Delta_{1,y}(\Delta_{2,z}\Delta_{3,x} - \Delta_{2,x}\Delta_{3,z})
+ \Delta_{1,z}(\Delta_{2,x}\Delta_{3,y} - \Delta_{2,y}\Delta_{3,x}) \;>\; 0.
\]
Expanding the $\Delta$'s as differences of original coordinates:
\[
\begin{aligned}
&(x_1{-}x_0)\bigl[(y_2{-}y_0)(z_3{-}z_0) - (z_2{-}z_0)(y_3{-}y_0)\bigr] \\
+\;&(y_1{-}y_0)\bigl[(z_2{-}z_0)(x_3{-}x_0) - (x_2{-}x_0)(z_3{-}z_0)\bigr] \\
+\;&(z_1{-}z_0)\bigl[(x_2{-}x_0)(y_3{-}y_0) - (y_2{-}y_0)(x_3{-}x_0)\bigr] \;>\; 0.
\end{aligned}
\]
Again the $S_4$-antisymmetry is hidden inside the differences.
\emph{Fully unparenthesized.} Distributing all subtractions yields the $24$-term
Levi-Civita expansion:
\[
[0,1,2,3] \;=\; \sum_{(d,a,b,c)\in S_4} \mathrm{sgn}(d,a,b,c)\, x_a\, y_b\, z_c,
\]
where $(d,a,b,c)$ ranges over permutations of $(0,1,2,3)$ and $d$ is the row
contributing the constant $1$ from the bordered determinant (so $d$ enters only
through the sign). Grouped by missing index $d$:
\[
\begin{aligned}
d{=}0:\quad &+\,x_1 y_2 z_3 - x_1 y_3 z_2 - x_2 y_1 z_3 + x_2 y_3 z_1 + x_3 y_1 z_2 - x_3 y_2 z_1 \\
d{=}1:\quad &-\,x_0 y_2 z_3 + x_0 y_3 z_2 + x_2 y_0 z_3 - x_2 y_3 z_0 - x_3 y_0 z_2 + x_3 y_2 z_0 \\
d{=}2:\quad &+\,x_0 y_1 z_3 - x_0 y_3 z_1 - x_1 y_0 z_3 + x_1 y_3 z_0 + x_3 y_0 z_1 - x_3 y_1 z_0 \\
d{=}3:\quad &-\,x_0 y_1 z_2 + x_0 y_2 z_1 + x_1 y_0 z_2 - x_1 y_2 z_0 - x_2 y_0 z_1 + x_2 y_1 z_0
\end{aligned}
\]
Every monomial uses each index at most once, and any swap of two indices flips the
signs of all $24$ terms in lock-step --- the four-fold antisymmetry on $\{0,1,2,3\}$
is entirely visible.

\section*{Condition 2: \texorpdfstring{$012^+ \cap 013^+ \neq \emptyset$}{012+ cap 013+ nonempty}}

Existence of $p$ with $[0,1,2,p] > 0$ \emph{and} $[0,1,3,p] > 0$.

\paragraph{(a) Determinant.}
\[
\exists\, p\in\mathbb{R}^3:\quad
\det\!\begin{pmatrix} 1 & x_0 & y_0 & z_0 \\ 1 & x_1 & y_1 & z_1 \\ 1 & x_2 & y_2 & z_2 \\ 1 & p_x & p_y & p_z \end{pmatrix} > 0
\;\wedge\;
\det\!\begin{pmatrix} 1 & x_0 & y_0 & z_0 \\ 1 & x_1 & y_1 & z_1 \\ 1 & x_3 & y_3 & z_3 \\ 1 & p_x & p_y & p_z \end{pmatrix} > 0
\]

\paragraph{(b) Scalar triple product.}
\[
\exists\, p\in\mathbb{R}^3:\quad
(P_1{-}P_0)\cdot\bigl((P_2{-}P_0)\times(p{-}P_0)\bigr) > 0
\;\wedge\;
(P_1{-}P_0)\cdot\bigl((P_3{-}P_0)\times(p{-}P_0)\bigr) > 0.
\]
\emph{Fully unparenthesized.} Each bracket cofactor-expands as in
Condition~1(b), with $p$ playing the role of the fourth point:
\[
\begin{aligned}
{[0,1,2,p]} \;&=\; P_1\!\cdot\!(P_2\!\times\!p) \;-\; P_0\!\cdot\!(P_2\!\times\!p) \;+\; P_0\!\cdot\!(P_1\!\times\!p) \;-\; P_0\!\cdot\!(P_1\!\times\!P_2),\\
{[0,1,3,p]} \;&=\; P_1\!\cdot\!(P_3\!\times\!p) \;-\; P_0\!\cdot\!(P_3\!\times\!p) \;+\; P_0\!\cdot\!(P_1\!\times\!p) \;-\; P_0\!\cdot\!(P_1\!\times\!P_3).
\end{aligned}
\]
The shared term $P_0\!\cdot\!(P_1\!\times\!p)$ reflects that both brackets share
the edge $\overline{P_0P_1}$. Each is $S_4$-antisymmetric in its own four points
($\{0,1,2,p\}$ resp.\ $\{0,1,3,p\}$) and \emph{affine} in $p$.

\paragraph{(c) Polynomial in coordinates.}
Substitute $p$ for $P_3$ in the first inequality, and $p$ for $P_2$ (keeping $P_3$
in the third slot) in the second. Writing $p=(x_p, y_p, z_p)$:
\[
\exists\,(x_p, y_p, z_p)\in\mathbb{R}^3:
\]
\[
\begin{aligned}
&(x_1{-}x_0)\bigl[(y_2{-}y_0)(z_p{-}z_0) - (z_2{-}z_0)(y_p{-}y_0)\bigr] \\
+\;&(y_1{-}y_0)\bigl[(z_2{-}z_0)(x_p{-}x_0) - (x_2{-}x_0)(z_p{-}z_0)\bigr] \\
+\;&(z_1{-}z_0)\bigl[(x_2{-}x_0)(y_p{-}y_0) - (y_2{-}y_0)(x_p{-}x_0)\bigr] \;>\; 0
\end{aligned}
\]
\[
\text{AND}
\]
\[
\begin{aligned}
&(x_1{-}x_0)\bigl[(y_3{-}y_0)(z_p{-}z_0) - (z_3{-}z_0)(y_p{-}y_0)\bigr] \\
+\;&(y_1{-}y_0)\bigl[(z_3{-}z_0)(x_p{-}x_0) - (x_3{-}x_0)(z_p{-}z_0)\bigr] \\
+\;&(z_1{-}z_0)\bigl[(x_3{-}x_0)(y_p{-}y_0) - (y_3{-}y_0)(x_p{-}x_0)\bigr] \;>\; 0.
\end{aligned}
\]
\emph{Fully unparenthesized.} Each bracket admits the $24$-term Levi-Civita
expansion of Condition~1(c). For $[0,1,2,p]$, treating $p$ as the fourth point
and grouping by missing index:
\[
\begin{aligned}
d{=}0:\quad &+\,x_1 y_2 z_p - x_1 y_p z_2 - x_2 y_1 z_p + x_2 y_p z_1 + x_p y_1 z_2 - x_p y_2 z_1 \\
d{=}1:\quad &-\,x_0 y_2 z_p + x_0 y_p z_2 + x_2 y_0 z_p - x_2 y_p z_0 - x_p y_0 z_2 + x_p y_2 z_0 \\
d{=}2:\quad &+\,x_0 y_1 z_p - x_0 y_p z_1 - x_1 y_0 z_p + x_1 y_p z_0 + x_p y_0 z_1 - x_p y_1 z_0 \\
d{=}p:\quad &-\,x_0 y_1 z_2 + x_0 y_2 z_1 + x_1 y_0 z_2 - x_1 y_2 z_0 - x_2 y_0 z_1 + x_2 y_1 z_0.
\end{aligned}
\]
$[0,1,3,p]$ is obtained by replacing every subscript $2$ with $3$ above. Two
structural observations are now visible:
\begin{itemize}
  \item Rows $d=0,1,2$ contain all the $p$-dependence (linear in $p$); row $d=p$
  is the constant $-\det(P_0,P_1,P_2)$ (resp.\ $-\det(P_0,P_1,P_3)$). So each
  bracket is affine in $p$, and the realizability question is whether two open
  halfspaces in $p$-space meet.
  \item The $S_4$-antisymmetry on the bracket's four points $\{0,1,2,p\}$ (or
  $\{0,1,3,p\}$) is manifest: every swap of two indices flips the signs of all
  $24$ monomials simultaneously.
\end{itemize}

\section*{Realizability: \texorpdfstring{$012^+\cap 013^+$}{012+ cap 013+} exists, \texorpdfstring{$012^-\cap 013^+\cap 123^+\cap 203^+$}{012- cap 013+ cap 123+ cap 203+} does not}

\paragraph{Notation conversion.} The bracket $[a,b,c,d]$ is alternating, so
$[2,0,3,p] = -[0,2,3,p]$, hence
\[
203^+ \;=\; \{p : [2,0,3,p] > 0\} \;=\; \{p : [0,2,3,p] < 0\} \;=\; 023^-.
\]
The cell in question is therefore $012^- \cap 013^+ \cap 023^- \cap 123^+$ in
canonical (sorted-index) notation --- the sign vector $(-,+,-,+)$ on the four
faces $(012, 013, 023, 123)$ of tetrahedron $0123$.

\paragraph{Key identity.} For any $p \in \mathbb{R}^3$,
\begin{equation}\label{eq:star}\tag{$\star$}
[0,1,2,p] \;-\; [0,1,3,p] \;+\; [0,2,3,p] \;-\; [1,2,3,p] \;=\; [0,1,2,3].
\end{equation}

\textit{Proof.} Since $P_0, P_1, P_2, P_3$ are affinely independent (because
$[0,1,2,3] \neq 0$), every $p \in \mathbb{R}^3$ has unique barycentric coordinates
\[
p \;=\; \alpha_0 P_0 + \alpha_1 P_1 + \alpha_2 P_2 + \alpha_3 P_3,
\qquad \alpha_0+\alpha_1+\alpha_2+\alpha_3 \;=\; 1.
\]
The lifted equality $(1,p) = \sum_i \alpha_i\,(1,P_i)$ holds in $\mathbb{R}^4$, so
multilinearity of the bordered $4\times 4$ determinant in row $p$ gives, for
distinct $a,b,c \in \{0,1,2,3\}$ with missing index $d$,
\[
[a,b,c,p] \;=\; \sum_{i=0}^{3} \alpha_i\, [a,b,c,i] \;=\; \alpha_d\, [a,b,c,d],
\]
since $[a,b,c,a] = [a,b,c,b] = [a,b,c,c] = 0$ (repeated row). Using
$[0,1,3,2] = -[0,1,2,3]$, $[0,2,3,1] = +[0,1,2,3]$, $[1,2,3,0] = -[0,1,2,3]$
(reorder the missing index back into canonical position), and dividing by
$[0,1,2,3]$:
\[
\alpha_3 \;=\; \frac{[0,1,2,p]}{[0,1,2,3]},\quad
\alpha_2 \;=\; \frac{-[0,1,3,p]}{[0,1,2,3]},\quad
\alpha_1 \;=\; \frac{[0,2,3,p]}{[0,1,2,3]},\quad
\alpha_0 \;=\; \frac{-[1,2,3,p]}{[0,1,2,3]}.
\]
Then $\alpha_0 + \alpha_1 + \alpha_2 + \alpha_3 = 1$, multiplied through by
$[0,1,2,3]$, is exactly~\eqref{eq:star}. \hfill$\square$

\paragraph{Why \texorpdfstring{$012^+ \cap 013^+$}{012+ cap 013+} is non-empty.}
Take the explicit witness
\[
p^\ast \;:=\; 2 P_3 - P_2 \qquad \text{(the reflection of } P_2 \text{ across } P_3\text{).}
\]
Its coefficients $(2, -1)$ sum to $1$, so any affine function of $p$ evaluated at
$p^\ast$ equals $2$ times its value at $P_3$ minus its value at $P_2$. With
$[0,1,2,P_2]=0$ and $[0,1,2,P_3]=[0,1,2,3]$:
\[
[0,1,2,\,p^\ast] \;=\; 2\,[0,1,2,3] \;-\; 0 \;=\; 2\,[0,1,2,3] \;>\; 0.
\]
With $[0,1,3,P_2]=[0,1,3,2]=-[0,1,2,3]$ and $[0,1,3,P_3]=0$:
\[
[0,1,3,\,p^\ast] \;=\; 2\cdot 0 \;-\; (-[0,1,2,3]) \;=\; [0,1,2,3] \;>\; 0.
\]
Both brackets are positive, so $p^\ast \in 012^+ \cap 013^+$. (Geometrically:
planes $012$ and $013$ are distinct and share the line $\overline{P_0 P_1}$, so
they cut $\mathbb{R}^3$ into four open wedges, all non-empty; $012^+\cap 013^+$
is one of them.)

\paragraph{Why \texorpdfstring{$012^- \cap 013^+ \cap 023^- \cap 123^+$}{012- cap 013+ cap 023- cap 123+} is empty.}
Suppose for contradiction such a $p$ exists. The four memberships translate to
\[
[0,1,2,p] < 0,\quad [0,1,3,p] > 0,\quad [0,2,3,p] < 0,\quad [1,2,3,p] > 0.
\]
Hence each summand on the left side of~\eqref{eq:star}, with its prescribed sign,
is \emph{strictly negative}:
\[
\underbrace{[0,1,2,p]}_{<\,0}
\;\underbrace{-\,[0,1,3,p]}_{<\,0}
\;\underbrace{+\,[0,2,3,p]}_{<\,0}
\;\underbrace{-\,[1,2,3,p]}_{<\,0}
\;<\; 0.
\]
But by~\eqref{eq:star} this sum equals $[0,1,2,3]$, which is $>0$ by Condition~1.
Contradiction. Hence $012^- \cap 013^+ \cap 023^- \cap 123^+ = \emptyset$.
\hfill$\square$

\paragraph{Geometric meaning.} In barycentric coordinates the four sign
conditions read
\[
\alpha_0 < 0,\quad \alpha_1 < 0,\quad \alpha_2 < 0,\quad \alpha_3 < 0
\qquad\text{(``$p$ on the wrong side of every face'')},
\]
which is impossible because $\alpha_0+\alpha_1+\alpha_2+\alpha_3 = 1 > 0$. The
four face-planes of $0123$ in general position cut $\mathbb{R}^3$ into $15$ open
cells (the count $\binom{4}{0}+\binom{4}{1}+\binom{4}{2}+\binom{4}{3}$), and the
missing $16^{\mathrm{th}}$ sign vector is precisely $(012^-, 013^+, 023^-, 123^+)$.
The asymmetry between the two cells of the question is now visible: $012^+\cap 013^+$
slices $\mathbb{R}^3$ along only \emph{two} planes (a wedge --- always non-empty),
while the missing $16^{\mathrm{th}}$ cell asks for a sign-coherent intersection
across \emph{all four} faces of a closed tetrahedron, which the affine relation
$\sum\alpha_i = 1$ rules out.

\section*{Observations}

\begin{itemize}
  \item \textbf{Only signs matter.} The cell complex depends only on the
  \emph{sign vector} in $\{+,-\}^{\binom{n}{3}}$. Coordinates are useful only as a
  realizability oracle: if we want a ``without coordinates'' enumeration, the
  coordinate forms are witnesses we consult only when the combinatorial sign rules
  cannot decide.

  \item \textbf{Condition 2 is automatic.} Planes $012$ and $013$ share the line
  $\overline{P_0P_1}$ and (since $P_3 \notin$ plane $012$ by Condition~1) are
  distinct. They partition $\mathbb{R}^3\setminus\overline{P_0P_1}$ into four open
  wedges, all non-empty. So $012^+\cap 013^+\neq\emptyset$ trivially. A concrete
  witness: any point $P_3 + \varepsilon\,\hat n_{013^+}$ with $\varepsilon$ small.

  \item \textbf{Where it gets interesting.} Once we have all four planes
  $\{012,\,013,\,023,\,123\}$ they cut $\mathbb{R}^3$ into
  $\binom{4}{0}+\binom{4}{1}+\binom{4}{2}+\binom{4}{3}=15$ regions. Of the $2^4=16$
  a-priori sign vectors, exactly one is empty --- the missing 16th cell. A clean
  way to spot it: the sign vector that says ``$p$ is on the \emph{opposite} side of
  every face from the opposite vertex of tetrahedron $0123$'' is not realizable
  (no point sees every face from the wrong side).
\end{itemize}

\section*{Algorithm: deciding if a new plane $ij(n{+}1)$ slices an existing cell}

We are now ready for the inductive step. Assume we have introduced $n+1$ named
points $\{P_0,\ldots,P_n\}$ and constructed the cell complex of the
$\binom{n+1}{3}$ planes through every triple. We pick a new point $P_{n+1}$
inside one specific cell $C^\ast$ of the current complex; this pins the signs of
all $4$-brackets involving $n{+}1$ (because membership of $P_{n+1}$ in $C^\ast$
is exactly $\sigma_{C^\ast}(\Pi) = \mathrm{sgn}([\Pi, n{+}1])$ for every existing
plane $\Pi$). Then we must slice every existing cell with the $\binom{n+1}{2}$
new planes $\bigl\{\,ij(n{+}1) : 0\le i<j\le n\,\bigr\}$. The basic question is:
given a cell $C$ and a new plane $\Pi = ij(n{+}1)$, does $\Pi$ slice $C$ (i.e.\
are both $C\cap\Pi^+$ and $C\cap\Pi^-$ non-empty)?

The algorithm below answers this question using the chirotope alone (no
coordinates), and is correct in the general case. The conceptual primitive is
the affine identity \eqref{eq:star} from earlier --- now applied to \emph{every}
$4$-tuple of named points, not just $\{0,1,2,3\}$.

\paragraph{State.} We maintain
\begin{itemize}
  \item the \textbf{chirotope}: $\mathrm{sgn}[a,b,c,d]$ for every $4$-subset
  $\{a,b,c,d\}\subseteq\{0,\ldots,n{+}1\}$;
  \item the \textbf{cell complex} of the previous step: each open cell $C$
  stored as a sign vector $\sigma_C : \mathcal{H}_{\mathrm{old}} \to \{+,-\}$
  over the old planes $\mathcal{H}_{\mathrm{old}} = \binom{\{0,\ldots,n\}}{3}$.
\end{itemize}
We do \emph{not} need anything else --- no coordinates, no magnitudes, no
``rays at infinity.''

\paragraph{The single sign-rule.} For every $4$-subset $\{a,b,c,d\}$ of named
indices and any point $p$, the affine identity \eqref{eq:star} (proved earlier
for $\{0,1,2,3\}$ but valid for any $4$-tuple, by the same barycentric
argument) reads
\[
  [a,b,c,p] \;-\; [a,b,d,p] \;+\; [a,c,d,p] \;-\; [b,c,d,p] \;=\; [a,b,c,d].
\]
Taking signs and noting the four magnitudes on the LHS are positive, the LHS
is forced to be strictly positive (regardless of magnitudes) iff the four
signs $(\sigma(abc), -\sigma(abd), \sigma(acd), -\sigma(bcd))$ are all $+$,
i.e.\ the cell sign pattern at the four ``faces of tetrahedron $abcd$'' is
\[
\bigl(\sigma(abc),\sigma(abd),\sigma(acd),\sigma(bcd)\bigr) \;=\; (+,-,+,-).
\]
Symmetrically the LHS is forced strictly negative iff the pattern is
$(-,+,-,+)$. So:

\noindent\textbf{Forbidden-pattern rule (per $4$-tuple).} A sign vector $\sigma$
violates the chirotope at $\{a,b,c,d\}$ iff
\[
\bigl(\sigma(abc),\sigma(abd),\sigma(acd),\sigma(bcd)\bigr) = \bigl(-,+,-,+\bigr)\cdot \mathrm{sgn}[a,b,c,d],
\]
where the right-hand side means: when $\mathrm{sgn}[a,b,c,d]=+$ the forbidden
pattern is $(-,+,-,+)$, and when $\mathrm{sgn}[a,b,c,d]=-$ the forbidden
pattern is $(+,-,+,-)$. This is exactly the ``missing $16^{\mathrm{th}}$ cell''
pattern from the original section, applied to every tetrahedron formed by four
named points.

\paragraph{Forbidden patterns are necessary, not sufficient.} A sign vector
$\sigma$ that corresponds to a non-empty cell of the realized arrangement
necessarily avoids every forbidden pattern --- this is what the affine
identity $(\star_{abcd})$ proves. The converse, however, fails once $N\ge
5$: there exist sign vectors that avoid every forbidden pattern yet are
not realized as cells. The first ``correction-tier'' is the
\emph{wedge-ordering} constraint, which we describe below.

\paragraph{Slice test (LP, with the forbidden-pattern rule as a pre-filter).}
Given an old cell $C$ with sign vector $\sigma_C$ and a new plane
$\Pi = ij(n{+}1)$, we ask in turn for each candidate sign $s\in\{+1,-1\}$:

\begin{enumerate}
  \item \emph{Cheap chirotope check.} For every $4$-subset
  $\{a,b,c,n{+}1\}$ such that $\Pi$ is one of its four faces (i.e.\
  $\{a,b,c\}\supseteq$ two of $\Pi$'s indices), see whether the forbidden
  pattern fires for $\sigma_C\cup\{\Pi\!\mapsto\!s\}$ on that 4-tuple.
  If yes, $C\cap\Pi^s$ is provably empty --- skip the LP.
  \item \emph{LP if not yet decided.} Otherwise, run an LP for feasibility
  of $\{\,\sigma\!\cdot\![\cdot,p]>0\,\}$ where $\sigma$ is $\sigma_C$
  extended by $\Pi\!\mapsto\!s$. If feasible, $C\cap\Pi^s$ is realized;
  otherwise it is not.
\end{enumerate}
The plane slices $C$ iff both halves are non-empty. The LP uses the named
vertices' coordinates (one specific realization of the chirotope-so-far),
not any cell witness; cell witnesses are not stored anywhere during
construction.

\paragraph{The Farkas analysis, and where the sufficiency argument fails.}
The cell $C\cap\Pi^+$ is non-empty iff the strict-linear system
\[
\bigl\{\,\sigma(abc)\,[a,b,c,p] > 0 : \text{all triples}\,\bigr\}
\]
is feasible in $p\in\mathbb{R}^3$. By Farkas, infeasibility requires a
non-negative combination $\sum \lambda_{abc}\,\sigma(abc)\,[a,b,c,p]$ that is
identically $\le 0$ as a function of $p$. Since $[a,b,c,p]$ is affine, this
means the linear part vanishes and the constant part is $\le 0$.

Differentiating $(\star_{abcd})$ in $p$ gives one family of linear-part-zero
combinations: the alternating $4$-term identity
\[
\mathbf{n}_{abc} \;-\; \mathbf{n}_{abd} \;+\; \mathbf{n}_{acd} \;-\; \mathbf{n}_{bcd} \;=\; 0
\]
(the dependency of plane normals around the tetrahedron $\{a,b,c,d\}$). Each
such dependency yields one Farkas certificate that fires exactly when its
forbidden pattern is hit, so the GP rule is \emph{necessary}.

But the tetrahedron-dependency is not the only $4$- or fewer-term
dependency. Consider the three planes $(i,j,a), (i,j,b), (i,j,c)$ that all
contain the line $\overline{P_iP_j}$: their normals are all perpendicular to
that line, hence lie in a $2$-plane, hence \emph{any three of them} are
linearly dependent. This is a $3$-term dependency among normals --- not the
alternating $4$-term one --- and it gives Farkas certificates that GP3 does
not see. Concretely, the $3$ planes through a common line cut a circular
arrangement of $6$ wedges around that line; only $6$ of the $2^3=8$ sign
patterns are realized, the other $2$ being forbidden. GP3 leaves all
$8$ admissible. This is the ``wedge-ordering'' constraint.

Higher-order Pl\"ucker relations contribute further constraints, and at
$n\ge 7$ M\"ne\v{v}--Sturmfels universality applies to general realizability
questions. So a complete chirotope-only characterization of topes is real
and hard.

\paragraph{What this means for the algorithm.} GP3 is a fast pre-filter:
when it fires, the extension is provably empty and we can skip the LP.
When it does not fire, we have to ask the LP (using the vertex
realizations as input data) for the final yes/no. The LP is exact and
chirotope-respecting --- different realizations of the same chirotope
give the same yes/no answer --- so the LP-based slicer enumerates the
correct cell complex. The vertex witnesses are needed by the LP only as a
specific representative of the chirotope; the cell complex shape is
intrinsic to the chirotope.

\paragraph{Why this is still not NP-complete in our setting.} Realizability
of an \emph{arbitrary} abstract chirotope is $\exists\mathbb{R}$-complete
(Mn\"ev), but that is not the question we ask: by construction we always
extend a chirotope that is already realized. For an already-realized
configuration, each slice test is just LP feasibility, which lies far
below the worst-case complexity wall.

\paragraph{Per-step complexity.} For step $n \to n{+}1$: $\binom{n+1}{2}$ new
planes, each tested against every old cell. The number of old cells is
bounded by Zaslavsky's count, $O(n^9)$. Each slice test is a single LP
feasibility check (or zero, when the GP forbidden-pattern pre-filter
fires), which is polynomial-time. The whole step is polynomial in $n$,
dominated by the cell count.

\section*{Segment-triangle piercing in bracket form}

A natural geometric primitive (and a check we will need later, e.g.\ for
``does this Cs\'asz\'ar candidate actually embed without crossings?''): given
two segments and a triangle made of named points, does the segment
$\overline{P_iP_j}$ pierce the (open) triangle $\triangle P_kP_lP_m$? The
answer is a four-line test in the chirotope, no coordinates needed.

\paragraph{Two conditions.} Assume the five points are in general position
(no $4$ coplanar, so all five $4$-brackets on $\{i,j,k,l,m\}$ are nonzero).
The segment pierces the triangle iff both:

\begin{enumerate}
  \item \emph{The line through $P_i,P_j$ pierces the triangle's interior},
  i.e.\ the three signed volumes of segment-against-edge tetrahedra all
  agree:
  \[
    \mathrm{sgn}[i,j,k,l] \;=\; \mathrm{sgn}[i,j,l,m] \;=\; \mathrm{sgn}[i,j,m,k]
    \;\in\;\{+,-\}.
  \]
  Each $[i,j,\cdot,\cdot]$ is, up to sign, the signed volume of the
  tetrahedron formed by the segment and one edge of the triangle. Their
  common sign records the directed orientation of $\overline{P_iP_j}$
  ``going around'' the triangle. If any one differs, the line misses or
  grazes that edge.

  \item \emph{The two endpoints $P_i,P_j$ lie on opposite sides of the
  plane through $klm$}, so the crossing point of the line with the plane
  falls inside the segment:
  \[
    \mathrm{sgn}[k,l,m,i] \;\neq\; \mathrm{sgn}[k,l,m,j].
  \]
\end{enumerate}

So the piercing test reads off five $4$-bracket signs on $\{i,j,k,l,m\}$ and
checks two equalities. The five brackets are tied by the GP identity
$\sum_q (-1)^q [\widehat q] = 0$, so any one is determined up to sign by the
other four; in practice we just take whichever signs the chirotope already
has.

\paragraph{Two-segment intersection.} As an aside: two segments
$\overline{P_iP_j}$ and $\overline{P_kP_l}$ in $\mathbb{R}^3$ can meet inside
their interiors only if they are coplanar, which requires $[i,j,k,l] = 0$
--- a non-generic condition. So in the generic chirotope two named segments
share at most a common endpoint, and $\mathrm{sgn}[i,j,k,l]$ records which
of the two segments ``goes above'' the other (the sign of the lift).

\section*{Redundant halfspaces in cell descriptions}

A cell of our arrangement is described by signs on \emph{every} plane, but
only some of these signs are \emph{essential}: dropping a non-essential one
leaves the cell unchanged. A non-essential constraint is \emph{redundant};
the planes whose constraints are essential are the \emph{facets} of the
cell.

\paragraph{The tetrahedron example.} In the four-plane arrangement of
tetrahedron $0123$, the cell ``past $P_3$'' (the open cone past vertex
$P_3$, bounded below by the three faces meeting at $P_3$) has sign vector
\[
  \bigl(\sigma(012),\sigma(013),\sigma(023),\sigma(123)\bigr) = (+,+,-,+).
\]
It really has only three facets --- the three faces $013, 023, 123$ that
meet at $P_3$. The fourth sign $\sigma(012) = +$ is redundant: the cell
happens to lie on the same side of $012$ as the tetrahedron's interior,
but the plane $012$ is not a face of the cell. Equivalently, flipping
$\sigma(012)$ to $-$ would give the sign vector $(-,+,-,+)$, which is
precisely the missing $16$th cell --- empty. So this constraint cannot be
the binding one.

\paragraph{The chirotope-only redundancy test.} The previous paragraph
generalizes: \emph{the constraint $\sigma_j$ on plane $\Pi_j = abc$ is
redundant for a cell $C$ iff flipping it produces an empty cell, iff some
forbidden-pattern rule is triggered after the flip}. Concretely,

\begin{quote}
$\sigma(abc)$ is redundant for $C$ iff there is some $d \notin \{a,b,c\}$
such that, after flipping $\sigma(abc)$, the four signs
$\bigl(\sigma(abc),\sigma(abd),\sigma(acd),\sigma(bcd)\bigr)$ form the
forbidden pattern $(-,+,-,+)\cdot\mathrm{sgn}[a,b,c,d]$ of the $4$-tuple
$\{a,b,c,d\}$.
\end{quote}

This is a pure chirotope check, with $O(n)$ work per plane (one
$4$-tuple per choice of $d$). It tells us, for each cell, exactly which
sign-constraints are facets and which are mere consequences of the rest.

\paragraph{Speeding up cell splitting.} The same machinery serves as a
pre-filter for the slice test. To decide whether a new plane $\Pi$ slices
cell $C$, we previously had to ask the LP about both extensions $\sigma^+,
\sigma^-$. With the chirotope filter:

\begin{enumerate}
  \item Run the forbidden-pattern check on $\sigma^+$ and on $\sigma^-$ in
  $O(n)$ time, considering only the $4$-tuples that involve $\Pi$'s three
  indices.
  \item If both $\sigma^+$ and $\sigma^-$ are forbidden-free: $\Pi$ slices
  $C$, do the actual split.
  \item If exactly one is forbidden: the cell lies entirely on the other
  side; attach that sign to $C$ and move on without an LP.
  \item Both being forbidden cannot happen for an existing cell --- $C$
  has a witness which already realizes one extension.
\end{enumerate}

In our $n=7$ run with $\sim\!3000$ cells, the new-plane stage at $n=8$ would
ask $\sim\!3000\times 21$ slice questions; the filter answers most of them
without any LP, since for most cells most new planes are redundant.

\paragraph{Storing only facets.} As a representational optimization, each
cell can be stored as the dict of its \emph{facet} signs (with the rest
re-derived from the chirotope on demand). With this convention, the
interior of tetrahedron $0123$ lists all four signs (four facets); the
past-$P_3$ cone lists only three; and the forbidden $16$th cell, if it were
realized, would list zero --- which is precisely the way to recognize that
it is empty.

\end{document}
